---
title: "Untitled"
author: "Kelvias Keter"
date: "2025-07-06"
output: html_document
editor_options: 
  chunk_output_type: console
---

## knitr options

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = FALSE,echo = TRUE)
```

### Load packages

```{r }
library(pacman)

p_load("tidyverse", "readxl","gtsummary","gt", "writexl", "ggpubr", "ggplot2","here","dgof", "reshape2","psych","DescTools","mdatools","lme4", "Matrix","sjPlot","rms","here","bstfun","patchwork","rstatix", "car","flextable","purrr","lattice","cutpointr","binom","devtools","effsize")

# devtools::install_github("MSKCC-Epi-Bio/bstfun")


```

## Set paths and load data

This code chunk aims to import the shigella dataset

```{r}
shigella_dataset<-read.csv("C:\\Users\\Kelvias Keter\\Downloads\\for_shigella_full_dataset(1).csv")

Flex_2a_dataset <- read.csv("C:\\Users\\Kelvias Keter\\Downloads\\Flex2a.csv")

```


### filter northern uganda
```{r}
Flex_2a_dataset <- Flex_2a_dataset %>%
  filter(site %in% c("Kilifi", "Kisumu", "Nairobi"))
```


## Renaming variables

```{r}

## rename participants in Flex_2a_dataset
Flex_2a_dataset <- Flex_2a_dataset %>% 
  rename(Sample_ID = pID.1)

###rename column2 in Flex_2a_dataset

Flex_2a_dataset <- Flex_2a_dataset %>% 
  rename(Flex_2a_IgG = Column2)

###rename sex.1 in Flex_2a_dataset
Flex_2a_dataset <- Flex_2a_dataset %>% 
  rename(sex = Sex.1)

###rename age in Flex_2a_dataset
Flex_2a_dataset <- Flex_2a_dataset %>% 
  rename(age_years = Age.1)

###rename age cat in Flex_2a_dataset
Flex_2a_dataset <- Flex_2a_dataset %>% 
  rename(age_cat = Sheet1.age_cat)

###rename study number in shigella_dataset
shigella_dataset <- shigella_dataset %>% 
  rename(Sample_ID = study_no)

```

###merge the two dataset
```{r}
## checking duplication in Flex_2a_dataset
dup <- sum(duplicated(Flex_2a_dataset$Sample_ID))  # Counts the number of duplicate rows
print(dup)

## checking duplication in shigella_dataset
dup <- sum(duplicated(shigella_dataset$Sample_ID))  # Counts the number of duplicate rows
print(dup)

# Using match() to find matches
matches <- shigella_dataset$Sample_ID %in% Flex_2a_dataset$Sample_ID
print(matches)  # Returns TRUE/FALSE for each participant

### checking if they match
matches <- match(shigella_dataset$Sample_ID, Flex_2a_dataset$Sample_ID)

print(matches)

# Count the number of matches
num_matches <- sum(!is.na(matches))
print(num_matches)

### dataset with IgG and sample ID
Flex_2a_dataset1 <- Flex_2a_dataset  %>% 
  select(c(Sample_ID,Flex_2a_IgG))

# Find matching participants
Merged_dataset <- inner_join(Flex_2a_dataset1, shigella_dataset, by = "Sample_ID")


## rename sex Merged_dataset
Merged_dataset$sex <- replace(Merged_dataset$sex,
                                   Merged_dataset$sex == "m", 
                                   "Male")
Merged_dataset$sex <- replace(Merged_dataset$sex,
                                   Merged_dataset$sex == "f", 
                                   "Female")

```

###creating age groups
```{r}

Merged_dataset <- Merged_dataset %>%
  mutate(Age_group = case_when(
    age_years < 5 ~ "<5 yrs",
    age_years >= 5 & age_years <= 9 ~ "5-9 yrs",
    age_years >= 10 & age_years <= 15 ~ "10-15 yrs",
    TRUE ~ NA_character_
  ))%>%
  relocate(Age_group, .after = age_years)
```

## creating a new variable of high and low responses
```{r}

# Calculate 66.7th percentile (upper tertile cutoff)
tertile_cutoff <- quantile(Merged_dataset$Flex_2a_IgG, probs = 2/3, na.rm = TRUE)

# Create tertile group: "Low" (bottom 2/3), "High" (top 1/3)
Merged_dataset <- Merged_dataset %>%
  mutate(Flex_2a_tertile = ifelse(Flex_2a_IgG <= tertile_cutoff, "Low", "High"))%>%
  relocate(Flex_2a_tertile, .after = Flex_2a_IgG)

```

### arranging variables
```{r}
library(dplyr)

Merged_dataset <- Merged_dataset %>%
  relocate(site, .before = Flex_2a_IgG)%>%
  relocate(sex, .before = site)%>%
  relocate(dob, .before = sex)%>%
  relocate(age_months, .before = sex)%>%
  relocate(age_years, .before = sex)%>%
  relocate(Age_group, .before = sex)%>%
  relocate(weight, .before = floor_type)%>%
  relocate(height, .before = floor_type)%>%
  relocate(muac, .before = floor_type)%>%
  relocate(site, .before = Sample_ID)
```

## creating variable ( Nutrition status by muac)
```{r}
Merged_dataset <- Merged_dataset %>%
  mutate(Nutrition_status = case_when(
    muac < 11.5 ~ "Severe acute malnutrition",
    muac >= 11.5 & muac < 12.5 ~ "Moderate acute malnutrition",
    muac >= 12.5~ "Normal",
    TRUE ~ NA_character_
  ))%>%
  relocate(Nutrition_status, .after = muac)
```

### creating variable (BMI)
```{r}
Merged_dataset <- Merged_dataset %>%
  mutate(BMI = weight / ( (height / 100)^2 ))%>%
  relocate(BMI, .before = muac)

```

###Creating age in days
```{r}
Merged_dataset <- Merged_dataset %>%
  mutate(Age_days= age_years*365)%>%
  relocate(Age_days, .before = age_years)
```

### duplicating sex to create coded sex
```{r}
Merged_dataset$coded_sex <- Merged_dataset$sex

###relocating the new column
Merged_dataset <- Merged_dataset %>%
  relocate(coded_sex, .after = sex)

###rename the observations in coded_sex
Merged_dataset$coded_sex <- replace(Merged_dataset$coded_sex,
                                   Merged_dataset$coded_sex == "Male", 
                                   "1")
Merged_dataset$coded_sex <- replace(Merged_dataset$coded_sex,
                                   Merged_dataset$coded_sex == "Female", 
                                   "2")


```

###calculating weight for height zscores
```{r}
library(zscorer)

Merged_dataset <- addWGSR(data = Merged_dataset, sex = "coded_sex", firstPart = "weight",
               secondPart = "height", index = "wfh")
###relocating the new column
Merged_dataset <- Merged_dataset %>%
  relocate(wfhz, .after = height)


```
###duplicating IgG and converting to a binary response
```{r}
Merged_dataset$Flex_2a_tertile_coded <- Merged_dataset$Flex_2a_tertile

###relocating the new column
Merged_dataset <- Merged_dataset %>%
  relocate(Flex_2a_tertile_coded, .after = Flex_2a_tertile)

###rename the observations in Flex_2a_tertile_coded
Merged_dataset$Flex_2a_tertile_coded <- replace(Merged_dataset$Flex_2a_tertile_coded,
                                   Merged_dataset$Flex_2a_tertile_coded == "Low", 
                                   "0")
Merged_dataset$Flex_2a_tertile_coded <- replace(Merged_dataset$Flex_2a_tertile_coded,
                                   Merged_dataset$Flex_2a_tertile_coded == "High", 
                                   "1")

```
### converting Flex_2a_tertile_coded as numeric
```{r}
Merged_dataset$Flex_2a_tertile_coded <- as.factor(Merged_dataset$Flex_2a_tertile_coded)

```


###regression for IgG high responses
```{r}
# Set a valid theme element
list("tbl_regression-str:ref_row_text" = "Ref") %>%
  set_gtsummary_theme() 

Merged_dataset$sex <- as.factor(Merged_dataset$sex)
Merged_dataset$Age_group <- as.factor(Merged_dataset$Age_group)
Merged_dataset$Nutrition_status <- factor( Merged_dataset$Nutrition_status,
  levels = c("Normal", "Moderate acute malnutrition", "Severe acute malnutrition")  # Put the reference first
)

## Univariate analysis
reg_model <- Merged_dataset %>%
  dplyr::select(Flex_2a_tertile_coded, site, Age_group, sex, Nutrition_status) %>% 
  tbl_uvregression(
    label = list(
      site ~ "Location", 
      Age_group ~ "Age (Yrs)", 
      sex ~ "Gender", 
      Nutrition_status ~ "Nutrition status (MUAC)"
    ),
    method = glm,
    y = Flex_2a_tertile_coded,
    method.args = list(family = "binomial"),
    exponentiate = TRUE
  ) %>% 
  bold_p(t = 0.05) %>% 
  modify_header(label = "**Covariates**") %>% 
  italicize_levels() 

reg_model

# Multivariate analysis
bm1 <- glm(Flex_2a_tertile_coded ~ site + Age_group + sex + Nutrition_status, 
           data = Merged_dataset, family = binomial)

bm_table1 <- tbl_regression(
    bm1,
    exponentiate = TRUE,
    label = list(
      site ~ "Location", 
      Age_group ~ "Age (Yrs)", 
      sex ~ "Gender", 
      Nutrition_status ~ "Nutrition status (MUAC)"
    )
  ) %>% 
  bold_p(t = 0.05) %>% 
  modify_header(label = "**Covariates**") %>% 
  italicize_levels() 

bm_table1

# Combine
uni_multi <- tbl_merge(
  tbls = list(reg_model, bm_table1), 
  tab_spanner = c("**Univariate analysis**", "**Multivariate analysis**")
)

uni_multi

# Save
uni_multi %>%
  as_flex_table() %>%
  save_as_image(path = "fancy_tableIgG_FLEX2a.png")

```
###roc curve
```{r}

library(pROC)

# Predict probabilities
prob <- predict(bm1, type = "response")

# Compute ROC curve
roc_obj <- roc(model_data$Flex_2a_tertile_coded, prob)

# Plot ROC curve
plot(roc_obj, col = "#1f78b4", lwd = 2, main = "ROC Curve")
abline(a = 0, b = 1, lty = 2, col = "gray")

# Calculate AUC
auc_value <- auc(roc_obj)
cat("AUC:", auc_value, "\n")

# Calculate 95% confidence interval for AUC
auc_ci <- ci.auc(roc_obj)
cat("95% CI for AUC:", auc_ci[1], "-", auc_ci[3], "\n")

# Optional: Add the CI text to the plot
legend("bottomright", legend = paste0("AUC = ", round(auc_value, 3), 
                                      "\n95% CI: ", round(auc_ci[1], 3), "-", round(auc_ci[3], 3)),
       bty = "n")


```

## Code to save a merged dataset

```{r}
write_csv(Merged_dataset, "Merged_dataset.csv")

```

## Participant characteristics 


```{r}

Merged_dataset%>% dplyr::select(c(sex,Age_group,site,Nutrition_status, Age_groups)) %>%
  tbl_summary(by=site,
    label = list(sex~"Gender",
                 Age_groups~"Age (Years)",
                 Nutrition_status~"Nutrition status"),
    missing = "no"
                 ) %>%
  add_stat_label(label=all_continuous()~"Median (IQR)") %>% 
  add_overall() %>% 
  add_variable_grouping ("Demographics"=c(
    "sex",
    "Age_groups")) %>%
  add_variable_grouping (" Nutrition status (MUAC)"=c(
    "Nutrition_status")) %>% 
  add_n() %>% 
  add_p() %>% 
  bold_labels() %>% 
  bold_p() %>% 
  modify_caption("Table 2. Participant characteristics") %>% 
  as_gt() %>% 
  gt::gtsave(filename = "tblxticsshigella.png",vheight = 1000)
  

```

###log transformation of Flex_2a_IgG
```{r}
Merged_dataset <- Merged_dataset %>% 
  mutate(Flex_2a_IgG_log=log(Flex_2a_IgG)) %>% 
  dplyr::relocate(Flex_2a_IgG_log, .after = Flex_2a_IgG)

```

### Boxplots for log-transformed IgG ELISA units by age 
```{r}
library(dplyr)
library(ggpubr)
library(ggplot2)

# Create Age_group variable with corrected labels
Merged_dataset <- Merged_dataset %>%
  mutate(
    Age_group = case_when(
      age_years < 5 ~ "<5 yrs",
      age_years >= 5 & age_years <= 9 ~ "5-9 Yrs",         # Corrected casing
      age_years >= 10 & age_years <= 15 ~ "10-15 Yrs",     # Corrected casing
      TRUE ~ NA_character_
    )
  ) %>%
  relocate(Age_group, .after = age_years)

# Convert to factor with correct level order
Merged_dataset$Age_group <- factor(Merged_dataset$Age_group,
                                   levels = c("<5 yrs", "5-9 Yrs", "10-15 Yrs"))

# Check distribution (optional)
table(Merged_dataset$Age_group, useNA = "ifany")

# Set method types
method1 <- "kruskal.test"   # or "anova"
method2 <- "wilcox.test"    # or "t.test"

# Comparisons for post-hoc tests
my_comparisations1 <- list(
  c("<5 yrs", "5-9 Yrs"),
  c("<5 yrs", "10-15 Yrs"),
  c("5-9 Yrs", "10-15 Yrs")
)

# Ensure complete data for plotting
Merged_dataset_plot <- Merged_dataset %>%
  filter(!is.na(Flex_2a_IgG_log), !is.na(Age_group))

# Plot with group comparisons
ggboxplot(Merged_dataset_plot,
          x = "Age_group", y = "Flex_2a_IgG_log",
          color = "Age_group", palette = "npg",
          add = "jitter", legend = "none") +
  stat_compare_means(
    aes(label = paste0(after_stat(method), ", p = ", after_stat(p.format))),
    method = method1,
    label.y = max(Merged_dataset_plot$Flex_2a_IgG_log, na.rm = TRUE, y.label=12) 
  ) +
  stat_compare_means(
    comparisons = my_comparisations1,
    method = method2,
    label = "p.signif"
  ) +
  labs(title = "", x = "Age", y = "Log-transformed IgG ELISA Units") +
  theme(text = element_text(size = 20))

folder_path <-"C:\\Users\\Kelvias Keter\\OneDrive\\Desktop\\msc project\\MSc Data analysis\\rplots"

ggsave("Log-transformed flex_IgG_byage.png", path=folder_path, dpi=800, width = 8, height = 6)


```

### Boxplots for log-transformed IgG ELISA units by location
```{r}

# Convert to factor with correct level order
Merged_dataset$site <- factor(Merged_dataset$site,
                                   levels = c("Kilifi", "Nairobi", "Kisumu"))

# Check distribution (optional)
table(Merged_dataset$Age_group, useNA = "ifany")

# Set method types
method1 <- "kruskal.test"   # or "anova"
method2 <- "wilcox.test"    # or "t.test"

# Comparisons for post-hoc tests
my_comparisations2 <- list(
  c("Nairobi", "Kilifi"),
  c("Nairobi", "Kisumu"),
  c("Kilifi", "Kisumu")
)

# Ensure complete data for plotting
Merged_dataset_plot <- Merged_dataset %>%
  filter(!is.na(Flex_2a_IgG_log), !is.na(site))

# Plot with group comparisons
ggboxplot(Merged_dataset_plot,
          x = "site", y = "Flex_2a_IgG_log",
          color = "site", palette = "npg",
          add = "jitter", legend = "none") +
  stat_compare_means(
    aes(label = paste0(after_stat(method), ", p = ", after_stat(p.format))),
    method = method1,
    label.y = max(Merged_dataset_plot$Flex_2a_IgG_log, na.rm = TRUE, y.label=15) 
  ) +
  stat_compare_means(
    comparisons = my_comparisations2,
    method = method2,
    label = "p.signif"
  ) +
  labs(title = "", x = "Location", y = "Log-transformed IgG ELISA Units") +
  theme(text = element_text(size = 20))

folder_path <-"C:\\Users\\Kelvias Keter\\OneDrive\\Desktop\\msc project\\MSc Data analysis\\rplots"

ggsave("Log-transformed flex_IgG_bylocation.png", path=folder_path, dpi=800, width = 8, height = 6)

```
### Boxplots for log-transformed IgG ELISA units by gender
```{r}

my_comparisations3 <- list( c("Female", "Male"))

x <- which(names(Merged_dataset) == "sex") # name of grouping variable
y <- which(
  names(Merged_dataset) == "Flex_2a_IgG_log" # names of variables to test
)

method <- "wilcox.test" # one of "wilcox.test" or "t.test"

Merged_dataset %>%
  select(c(Flex_2a_IgG_log, sex))


for (i in y) {
  for (j in x) {
    p <- ggboxplot(Merged_dataset,
      x = colnames(Merged_dataset[j]), y = colnames(Merged_dataset[i]),
      color = colnames(Merged_dataset[j]),
      legend = "none",
      palette = "npg",
      add = "jitter"
    )
    print(
      p + stat_compare_means(aes(label = paste0(after_stat(method), ", p-value = ", after_stat(p.format))),
        method = method, label.y = max(Merged_dataset[, i], na.rm = TRUE, label.y = 15)
      )
      + stat_compare_means(comparisons =my_comparisations3 , method = method, label = "p.signif") # remove if p-value of ANOVA or Kruskal-Wallis test >= alpha 
       +
        labs(title = "",
       x = "Gender", y = "Log-transformed IgG ELISA Units") +
        theme(text = element_text(size=20))
    )
  }
}

folder_path <-"C:\\Users\\Kelvias Keter\\OneDrive\\Desktop\\msc project\\MSc Data analysis\\rplots"

ggsave("Log-transformed flex_IgG_bygender.png", path=folder_path, dpi=800, width = 8, height = 6)

```

### collapsing age to form < &> 5yrs
```{r}
Merged_dataset <- Merged_dataset %>%
  mutate(Age_groups = case_when(
    age_years < 5 ~ "<5 Yrs",
    age_years >= 5 ~ ">5 Yrs",
    TRUE ~ NA_character_
  ))%>%
  relocate(Age_groups, .after = age_years)
```

### Boxplots for log-transformed IgG ELISA units by age_groups
```{r}
library(dplyr)
library(ggpubr)
library(ggplot2)

# Recode new age group variable (if needed)
Merged_dataset <- Merged_dataset %>%
  mutate(
    Age_groups = case_when(
      age_years < 5 ~ "<5 Yrs",
      age_years >= 5 ~ ">5 Yrs",
      TRUE ~ NA_character_
    )
  )

# Convert to factor with correct level order
Merged_dataset$Age_groups <- factor(Merged_dataset$Age_groups,
                                    levels = c("<5 Yrs", ">5 Yrs"))

# Check distribution (optional)
table(Merged_dataset$Age_groups, useNA = "ifany")

# Set method types
method1 <- "kruskal.test"
method2 <- "wilcox.test"

# Comparison list must match factor levels
my_comparisations4 <- list(c("<5 Yrs", ">5 Yrs"))

# Filter complete cases for plotting
Merged_dataset_plot <- Merged_dataset %>%
  filter(!is.na(Flex_2a_IgG_log), !is.na(Age_groups))

# Check if data is non-empty before plotting
if (nrow(Merged_dataset_plot) > 0) {
  # Store plot in an object
  p <- ggboxplot(Merged_dataset_plot,
                 x = "Age_groups", y = "Flex_2a_IgG_log",
                 color = "Age_groups", palette = "npg",
                 add = "jitter", legend = "none") +
    stat_compare_means(
      aes(label = paste0(after_stat(method), ", p = ", after_stat(p.format))),
      method = method1,
      label.y = max(Merged_dataset_plot$Flex_2a_IgG_log, na.rm = TRUE) + 1
    ) +
    stat_compare_means(
      comparisons = my_comparisations4,
      method = method2,
      label = "p.signif"
    ) +
    labs(title = "", x = "Age", y = "Log-transformed IgG ELISA Units") +
    theme(text = element_text(size = 20))

  # Print plot
  print(p)

  # Save plot
  folder_path <- "C:\\Users\\Kelvias Keter\\OneDrive\\Desktop\\msc project\\MSc Data analysis\\rplots"
  ggsave("Log-transformed flex_IgG_by_under5_age.png", plot = p, path = folder_path, dpi = 800, width = 8, height = 6)
} 


```

## Participants categorized by age groups


```{r}
Merged_dataset <- Merged_dataset %>%
  mutate(
    Age_groups = case_when(
      age_years < 5 ~ "<5 Yrs",
      age_years >= 5 ~ ">5 Yrs",
      TRUE ~ NA_character_
    ),
    Age_groups = as.character(Age_groups, levels = c("<5 Yrs", ">5 Yrs"))  # explicitly categorical
  )

library(gtsummary)
library(gt)
library(dplyr)

summary_table <- Merged_dataset %>%
  select(sex, Age_groups, site, Nutrition_status, Flex_2a_IgG) %>%
  mutate(Flex_2a_IgG = as.numeric(Flex_2a_IgG)) %>%
  tbl_summary(
    by = Age_groups,
    label = list(
      sex ~ "Gender",
      site ~ "Location",
      Nutrition_status ~ "Nutrition status (MUAC)",
      Flex_2a_IgG ~ "Flex 2A IgG response"
    ),
    type = list(
      Flex_2a_IgG ~ "continuous"
    ),
    statistic = all_continuous() ~ "{median} ({p25}, {p75})",
    missing = "no"
  ) %>%
  add_overall() %>%
  add_n() %>%
  add_p() %>%
  bold_labels() %>%
  bold_p() %>%
  modify_caption("**Table 3. Participant characteristics**")

gtsave(
  as_gt(summary_table),
  filename = "tblxticsshigella_1.png",
  path = "C:/Users/Kelvias Keter/OneDrive/Desktop/msc project/MSc Data analysis/rplots",
  vheight = 1000
)



```

